---
title: "Results paper 1"
author: "Alejandro Muñoz Escribano"
format: html
editor: visual
---

```{r setup, include=FALSE}

library(tidyverse)
library(glmmTMB)
library(emmeans)
library(DHARMa)
library(performance)
library(car)
here::i_am("docs/Results.qmd")
```

# Lisso abundance

## With field 5

```{r Lisso dataframe, include=FALSE}
Lisso <- read.csv2(here::here("data/modified/Lisso_abundance.csv")) %>% 
  mutate(Date = as.factor(Date),
         Field = as.factor(Field),
         Treatment = as.factor(Treatment))

model.lisso <- glmmTMB(Abundance ~ Treatment*Date + (1|Field), data = Lisso, family = poisson)
```

**Model formula:**

```{r Lisso model, output=FALSE}
glmmTMB(Abundance ~ Treatment*Date + (1|Field), data = Lisso, family = poisson)
```

**Summary:**

```{r Lisso summary, echo=FALSE}
summary(model.lisso)
r2(model.lisso)
Anova(model.lisso, type = 3)
```

**Model diagnostics:**

```{r Lisso model diagnostics, echo=FALSE}
dharma <- simulateResiduals(model.lisso, plot = T)
```

**Estimated marginal means plots:**

Factor treatment

```{r Lisso emmeans 1, echo=FALSE, message=FALSE}
emmeans(model.lisso, pairwise~Treatment, type = "response")

lisso_emm <- as.data.frame(emmeans(model.lisso, pairwise ~ Treatment, type = "response")$emmeans)
ggplot(lisso_emm, aes(x = Treatment, y = rate)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = rate-SE, ymax = rate+SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of Adult Abundance by Treatment",
       x = "Treatment", y = "Abundance") +
  theme_minimal()
```

Factor treatment\*date

```{r Lisso emmeans 2, echo=FALSE, message=FALSE}
emmeans(model.lisso, pairwise~Treatment|Date, type = "response")

lisso_emm.2 <- as.data.frame(emmeans(model.lisso, pairwise ~ Treatment|Date, type = "response")$emmeans)
ggplot(lisso_emm.2, aes(x = Date, y = rate, color = Treatment)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = rate-SE, ymax = rate+SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of Adult Abundance by Treatment and Date",
       x = "Date", y = "Abundance") +  
  theme_minimal()
```

**Figure:**

```{r Lisso figure treatment, echo=FALSE}
pd_jitter <- position_jitter(width = 0.07, height = 0)
pd_dodge  <- position_dodge(width = 0.15)
pal_cb <- c(
  "#0072B2", # azul
  "#D55E00", # rojo/naranja
  "#009E73", # verde
  "#CC79A7", # morado
  "#E69F00"  # amarillo/naranja
)

lisso_summ <- Lisso %>% 
  group_by(Field, Treatment) %>% 
  summarise(mean_dmg = mean(Abundance),
            sd = sd(Abundance, na.rm = TRUE),
            n = sum(!is.na(Abundance)),
            se = sd / sqrt(n),
            .groups = "drop"
  )

#Plot
ggplot() +
  geom_jitter(
    data = Lisso,
    aes(x = Treatment, y = Abundance, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = lisso_summ,
    aes(x = Treatment, y = mean_dmg, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = lisso_summ,
    aes(x = Treatment, y = mean_dmg, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  #Global emmean layer
  geom_line(
    data = lisso_emm,
    aes(x = Treatment, y = rate, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = lisso_emm,
    aes(x = Treatment,
        ymin = rate-SE,
        ymax = rate+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = lisso_emm,
    aes(x = Treatment, y = rate),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Adult abundance",
  ) +
  theme_minimal(base_size = 14)
```

Factor treatment\*date

```{r Lisso figure date, echo=FALSE}
lisso_summ_date <- Lisso %>% 
  group_by(Field, Treatment, Date) %>% 
  summarise(mean_dmg = mean(Abundance),
            sd = sd(Abundance, na.rm = TRUE),
            n = sum(!is.na(Abundance)),
            se = sd / sqrt(n),
            .groups = "drop"
  )


ggplot() +
  geom_jitter(
    data = Lisso,
    aes(x = Treatment, y = Abundance, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = lisso_summ_date,
    aes(x = Treatment, y = mean_dmg, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = lisso_summ_date,
    aes(x = Treatment, y = mean_dmg, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  facet_wrap(~Date) +
  
  
  #Global emmean layer
  
  geom_line(
    data = lisso_emm.2,
    aes(x = Treatment, y = rate, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = lisso_emm.2,
    aes(x = Treatment,
        ymin = rate-SE,
        ymax = rate+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = lisso_emm.2,
    aes(x = Treatment, y = rate),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Abundancia de adultos",
  ) +
  theme_minimal(base_size = 14)
```



## Without field 5

```{r Lisso dataframe no-5, include=FALSE}
Lisso_no5<- Lisso %>% 
  filter (Field != "5")

model.lisso.no5 <- glmmTMB(Abundance ~ Treatment*Date + (1|Field), data = Lisso_no5, family = poisson)

```

**Summary:**

```{r Lisso summary no-5, echo=FALSE}
summary(model.lisso.no5)

r2(model.lisso.no5)

Anova(model.lisso.no5, type = 3)
```

**Model diagnostics:**

```{r Lisso model diagnostics no-5, echo=FALSE}
dharma <- simulateResiduals(model.lisso.no5, plot = T)
```

**Estimated marginal means plots:**

Factor treatment

```{r Lisso emmeans 1 no-5, echo=FALSE, message=FALSE}
emmeans(model.lisso.no5, pairwise~Treatment, type = "response")

lisso.no5_emm <- as.data.frame(emmeans(model.lisso.no5, pairwise ~ Treatment, type = "response")$emmeans)
ggplot(lisso.no5_emm, aes(x = Treatment, y = rate)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = rate-SE, ymax = rate+SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of Adult Abundance by Treatment",
       x = "Treatment", y = "Abundance") +
  theme_minimal()
```

Factor treatment\*date

```{r Lisso emmeans 2 no-5, echo=FALSE, message=FALSE}
emmeans(model.lisso.no5, pairwise~Treatment|Date, type = "response")

lisso.no5_emm.2 <- as.data.frame(emmeans(model.lisso.no5, pairwise ~ Treatment|Date, type = "response")$emmeans)
ggplot(lisso.no5_emm.2, aes(x = Date, y = rate, color = Treatment)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = rate-SE, ymax = rate+SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of Adult Abundance by Treatment and Date",
       x = "Date", y = "Abundance") +
  theme_minimal()
```

**Figure:**

Factor treatment

```{r Lisso figure treatment no-5, echo=FALSE}

lisso.no5_summ <- Lisso_no5 %>% 
  group_by(Field, Treatment) %>% 
  summarise(mean_dmg = mean(Abundance),
            sd = sd(Abundance, na.rm = TRUE),
            n = sum(!is.na(Abundance)),
            se = sd / sqrt(n),
            .groups = "drop"
  )

#Plot
ggplot() +
  geom_jitter(
    data = Lisso_no5,
    aes(x = Treatment, y = Abundance, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = lisso.no5_summ,
    aes(x = Treatment, y = mean_dmg, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = lisso.no5_summ,
    aes(x = Treatment, y = mean_dmg, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  #Global emmean layer
  geom_line(
    data = lisso.no5_emm,
    aes(x = Treatment, y = rate, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = lisso.no5_emm,
    aes(x = Treatment,
        ymin = rate-SE,
        ymax = rate+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = lisso.no5_emm,
    aes(x = Treatment, y = rate),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Adult abundance",
  ) +
  theme_minimal(base_size = 14)
```

Factor treatment\*date

```{r Lisso figure date no-5, echo=FALSE}

lisso.no5_summ_date <- Lisso_no5 %>% 
  group_by(Field, Treatment, Date) %>% 
  summarise(mean_dmg = mean(Abundance),
            sd = sd(Abundance, na.rm = TRUE),
            n = sum(!is.na(Abundance)),
            se = sd / sqrt(n),
            .groups = "drop"
  )

#Plot
ggplot() +
  geom_jitter(
    data = Lisso_no5,
    aes(x = Treatment, y = Abundance, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = lisso.no5_summ_date,
    aes(x = Treatment, y = mean_dmg, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = lisso.no5_summ_date,
    aes(x = Treatment, y = mean_dmg, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  facet_wrap(~Date) +
  
  #Global emmean layer
  geom_line(
    data = lisso.no5_emm.2,
    aes(x = Treatment, y = rate, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = lisso.no5_emm.2,
    aes(x = Treatment,
        ymin = rate-SE,
        ymax = rate+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = lisso.no5_emm.2,
    aes(x = Treatment, y = rate),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Adult abundance",
  ) +
  theme_minimal(base_size = 14)
```

------------------------------------------------------------------------



# Leaf damage

## With field 5

```{r Leaf dataframe, include=FALSE}
leaf_dmg <- read.csv2(here::here("data/original/leaf_dmg.csv")) %>% 
  mutate(Date = ifelse(Date == "2025-06-13", "2025-06-11", as.character(Date))) %>%
  mutate(Date = as.factor(Date),
         Field = as.character(Field), 
         Repeat = as.character(Repeat)) %>% 
  mutate(No_dmg = 10 - Leaves_dmg_10leaves) %>% 
  filter(Treatment %in% c("BE", "FO"))

model.leaf <- glmmTMB(cbind(Leaves_dmg_10leaves, No_dmg) ~ Treatment * Date + (1|Field), data = leaf_dmg, family = betabinomial)
```

**Model formula:**

```{r Leaf model, output=FALSE}
glmmTMB(cbind(Leaves_dmg_10leaves, No_dmg) ~ Treatment * Date + (1|Field), data = leaf_dmg, family = betabinomial)
```

**Summary:**

```{r Leaf summary, echo=FALSE}
summary(model.leaf)

r2(model.leaf)

Anova(model.leaf, type = 3)
```

**Model diagnostics:**

```{r Leaf model diagnostics, echo=FALSE}
dharma <- simulateResiduals(model.leaf, plot = T)
```

**Estimated marginal means plots:**

Factor treatment

```{r Leaf emmeans, echo=FALSE, message=FALSE}
emmeans(model.leaf, pairwise~Treatment, type = "response")

leaf_emm <- as.data.frame(emmeans(model.leaf, pairwise ~ Treatment, type = "response")$emmeans)
ggplot(leaf_emm, aes(x = Treatment, y = prob)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = prob-SE, ymax = prob+SE), width = 0.2) +
  labs(title = "Estimated Probability of Leaf Damage by Treatment",
       x = "Treatment",
       y = "Probability") + 
  theme_minimal()
```

**Figure:**

Factor treatment:

```{r Leaf figure treatment, echo = FALSE}
leaf_summ <- leaf_dmg %>% 
  group_by(Field, Treatment) %>% 
  summarise(mean_dmg = mean(Leaves_dmg_10leaves/10),
            sd = sd(Leaves_dmg_10leaves/10, na.rm = TRUE),
            n = sum(!is.na(Leaves_dmg_10leaves/10)),
            se = sd / sqrt(n),
            .groups = "drop"
  )

ggplot() +
  geom_jitter(
    data = leaf_dmg,
    aes(x = Treatment, y = Leaves_dmg_10leaves/10, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = leaf_summ,
    aes(x = Treatment, y = mean_dmg, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = leaf_summ,
    aes(x = Treatment, y = mean_dmg, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  
#Global emmean layer
  
  geom_line(
    data = leaf_emm,
    aes(x = Treatment, y = prob, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = leaf_emm,
    aes(x = Treatment,
        ymin = prob-SE,
        ymax = prob+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = leaf_emm,
    aes(x = Treatment, y = prob),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Damaged leaves proportion",
  ) +
  theme_minimal(base_size = 14)
```



## Without field 5

```{r Leaf dataframe no-5, include=FALSE}
leaf_dmg_no5 <- leaf_dmg %>% filter(Field != 5)

model.leaf.no5 <- glmmTMB(cbind(Leaves_dmg_10leaves, No_dmg) ~ Treatment * Date + (1|Field), data = leaf_dmg_no5, family = betabinomial)
```

**Summary:**

```{r Leaf summary no-5, echo=FALSE}
summary(model.leaf.no5)

r2(model.leaf.no5)

Anova(model.leaf.no5, type = 3)
```

**Model diagnostics:**

```{r Leaf model diagnostics no-5, echo=FALSE}
dharma <- simulateResiduals(model.leaf.no5, plot = T)
```

**Estimated marginal means plots:**

Factor treatment

```{r Leaf emmeans 1 no-5, echo=FALSE, message=FALSE}
emmeans(model.leaf.no5, pairwise~Treatment, type = "response")

leaf.no5_emm1 <- as.data.frame(emmeans(model.leaf.no5, pairwise ~ Treatment, type = "response")$emmeans)
ggplot(leaf.no5_emm1, aes(x = Treatment, y = prob)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = prob-SE, ymax = prob+SE), width = 0.2) +
  labs(title = "Estimated Probability of Leaf Damage by Treatment",
       x = "Treatment",
       y = "Probability") + 
  theme_minimal()
```

Factor treatment*date

```{r Leaf emmeans 2 no-5, echo=FALSE, message=FALSE}
emmeans(model.leaf.no5, pairwise~Treatment|Date, type = "response")

leaf.no5_emm2 <- as.data.frame(emmeans(model.leaf.no5, pairwise ~ Treatment|Date, type = "response")$emmeans)
ggplot(leaf.no5_emm2, aes(x = Date, y = prob, color = Treatment)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = prob-SE, ymax = prob+SE), width = 0.2) +
  labs(title = "Estimated Probability of Leaf Damage by Treatment and Date",
       x = "Date",
       y = "Probability") + 
  theme_minimal()
```

**Figure:**

Factor treatment:

```{r Leaf figure treatment no-5, echo = FALSE}
leaf.no5_summ <- leaf_dmg_no5 %>% 
  group_by(Field, Treatment) %>% 
  summarise(mean_dmg = mean(Leaves_dmg_10leaves/10),
            sd = sd(Leaves_dmg_10leaves/10, na.rm = TRUE),
            n = sum(!is.na(Leaves_dmg_10leaves/10)),
            se = sd / sqrt(n),
            .groups = "drop"
  )

ggplot() +
  geom_jitter(
    data = leaf_dmg_no5,
    aes(x = Treatment, y = Leaves_dmg_10leaves/10, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = leaf.no5_summ,
    aes(x = Treatment, y = mean_dmg, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = leaf.no5_summ,
    aes(x = Treatment, y = mean_dmg, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  
#Global emmean layer
  
  geom_line(
    data = leaf.no5_emm1,
    aes(x = Treatment, y = prob, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = leaf.no5_emm1,
    aes(x = Treatment,
        ymin = prob-SE,
        ymax = prob+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = leaf.no5_emm1,
    aes(x = Treatment, y = prob),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Damaged leaves proportion",
  ) +
  theme_minimal(base_size = 14)
```

Factor treatment*date

```{r Leaf figure date no-5, echo = FALSE}
leaf.no5_summ_date <- leaf_dmg_no5 %>% 
  group_by(Field, Treatment,Date) %>% 
  summarise(mean_dmg = mean(Leaves_dmg_10leaves/10),
            sd = sd(Leaves_dmg_10leaves/10, na.rm = TRUE),
            n = sum(!is.na(Leaves_dmg_10leaves/10)),
            se = sd / sqrt(n),
            .groups = "drop"
  )


ggplot() +
  geom_jitter(
    data = leaf_dmg_no5,
    aes(x = Treatment, y = Leaves_dmg_10leaves/10, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = leaf.no5_summ_date,
    aes(x = Treatment, y = mean_dmg, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = leaf.no5_summ_date,
    aes(x = Treatment, y = mean_dmg, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  
  #Global emmean layer
  
  geom_line(
    data = leaf.no5_emm2,
    aes(x = Treatment, y = prob, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = leaf.no5_emm2,
    aes(x = Treatment,
        ymin = prob-SE,
        ymax = prob+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = leaf.no5_emm2,
    aes(x = Treatment, y = prob),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Damaged leaves proportion",
  ) +
  theme_minimal(base_size = 14) +
  facet_wrap(~Date)
```

------------------------------------------------------------------------



# Root biomass

## With field 5

```{r Root dataframe, include=FALSE}
root <- read.csv2(here::here("data/original/Root_dmg_tot.csv")) %>% 
        mutate(Date = as.POSIXct(Date),
               Field = as.factor(Field),
               Repeats = as.character(Repeats)) %>% 
        filter(Treatment %in% c("BE", "FO"),
               Date == "2025-06-30",
               Root_weight < 3) %>% #Remove one outlier with weight value higher than 6
        mutate(Treatment = as.factor(Treatment)) %>% 
        select(-Observations)

model.root <- glmmTMB(Root_weight ~ Treatment + (1|Field), data = root, family = Gamma (link = "log"))
```

**Model formula:**

```{r Root model, output=FALSE}
glmmTMB(Root_weight ~ Treatment + (1|Field), data = root, family = Gamma (link = "log"))
```

**Summary:**

```{r Root summary, echo = FALSE}
summary(model.root)

r2(model.root)

Anova(model.root, type = 2)
```

**Model diagnostics:**

```{r Root model diagnostics, echo=FALSE}
dharma <- simulateResiduals(model.root, plot = T)
```

**Estimated marginal means plots:**

```{r Root emmeans, echo=FALSE, message=FALSE}
emmeans(model.root, pairwise~Treatment, type = "response")

root_emm <- as.data.frame(emmeans(model.root, pairwise ~ Treatment, type = "response")$emmeans)
ggplot(root_emm, aes(x = Treatment, y = response)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = response-SE, ymax = response+SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of Root Biomass by Treatment",
       x = "Treatment",
       y = "Root biomass (g)") + 
  theme_minimal()
```

**Figure:**

```{r Root figure, echo = FALSE}
root_summ <- root %>% 
  group_by(Field, Treatment) %>% 
  summarise(mean_weight = mean(Root_weight),
            sd = sd(Root_weight, na.rm = TRUE),
            n = sum(!is.na(Root_weight)),
            se = sd / sqrt(n),
            .groups = "drop"
  )

ggplot() +
  geom_jitter(
    data = root,
    aes(x = Treatment, y = Root_weight, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = root_summ,
    aes(x = Treatment, y = mean_weight, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = root_summ,
    aes(x = Treatment, y = mean_weight, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  
  #Global emmean layer
  
  geom_line(
    data = root_emm,
    aes(x = Treatment, y = response, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = root_emm,
    aes(x = Treatment,
        ymin = response-SE,
        ymax = response+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = root_emm,
    aes(x = Treatment, y = response),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Root biomass (g)",
  ) +
  theme_minimal(base_size = 14)
```



## Without field 5

```{r Root dataframe no-5, include=FALSE}
root_no5 <- root %>% filter(Field != "5")

model.root.no5 <- glmmTMB(Root_weight ~ Treatment + (1|Field), data = root_no5, family = Gamma (link = "log"))
```

**Summary:**

```{r Root summary no-5, echo = FALSE}
summary(model.root.no5)

r2(model.root.no5)

Anova(model.root.no5, type = 2)
```

**Model diagnostics:**

```{r Root model diagnostics no-5, echo=FALSE}
dharma <- simulateResiduals(model.root.no5, plot = T)
```

**Estimated marginal means plots:**

```{r Root emmeans no-5, echo=FALSE, message=FALSE}
emmeans(model.root.no5, pairwise~Treatment, type = "response")

root.no5_emm <- as.data.frame(emmeans(model.root.no5, pairwise ~ Treatment, type = "response")$emmeans)
ggplot(root.no5_emm, aes(x = Treatment, y = response)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = response-SE, ymax = response+SE), width = 0.2) +
  labs(title = "Estimated Marginal Means of Root Biomass by Treatment",
       x = "Treatment",
       y = "Root biomass (g)") + 
  theme_minimal()
```

**Figure:**

```{r Root figure no-5, echo = FALSE}
root.no5_summ <- root_no5 %>% 
  group_by(Field, Treatment) %>% 
  summarise(mean_weight = mean(Root_weight),
            sd = sd(Root_weight, na.rm = TRUE),
            n = sum(!is.na(Root_weight)),
            se = sd / sqrt(n),
            .groups = "drop"
  )


ggplot() +
  geom_jitter(
    data = root_no5,
    aes(x = Treatment, y = Root_weight, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = root.no5_summ,
    aes(x = Treatment, y = mean_weight, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = root.no5_summ,
    aes(x = Treatment, y = mean_weight, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  
  #Global emmean layer
  
  geom_line(
    data = root.no5_emm,
    aes(x = Treatment, y = response, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = root.no5_emm,
    aes(x = Treatment,
        ymin = response-SE,
        ymax = response+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = root.no5_emm,
    aes(x = Treatment, y = response),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Root biomass (g)",
  ) +
  theme_minimal(base_size = 14)
```

------------------------------------------------------------------------



# Yield

## With field 5

```{r Yield dataframe, include=FALSE}
yield <- read_csv2(here::here("data/original/Yield.csv")) %>% 
  mutate(Date = as.POSIXct(Date),
         Field = as.factor(Field), 
         Repeat = as.character(Repeat)) %>% 
  rename(Yield = `Yield_kg/ha_HR14`) %>%
  filter(Treatment %in% c("BE", "FO"))

model.yield <- glmmTMB(Yield ~ Treatment + (1|Field), data = yield, family = Gamma (link = "log"))
```

**Model formula:**

```{r Yield model, output=FALSE}
glmmTMB(Yield ~ Treatment + (1|Field), data = yield, family = Gamma (link = "log"))
```

**Summary:**

```{r Yield summary, echo = FALSE}
summary(model.yield)

r2(model.yield)

Anova(model.yield)
```

**Model diagnostics:**

```{r Yield model diagnostics, echo=FALSE}
dharma <- simulateResiduals(model.yield, plot = T)
```

**Estimated marginal means plots:**

```{r Yield emmeans, echo = FALSE, message = FALSE}
emmeans(model.yield, pairwise ~ Treatment, type = "response")

yield_emm <- as.data.frame(emmeans(model.yield, pairwise ~Treatment, type = "response")$emmeans)
ggplot(yield_emm, aes(x = Treatment, y = response))+
         geom_point(size = 2)+
         geom_errorbar(aes(ymin = response-SE, ymax = response+SE, width = 0.2))+
  labs(title = "Estimated Marginal Means of Yield by Treatment",
       x = "Treatment",
       y = "Yield (kg/ha)") +
  theme_minimal()
```

**Figure:**

```{r Figure yield, echo = FALSE}
yield_summ <- yield %>% 
  group_by(Field, Treatment) %>% 
  summarise(mean_yield = mean(Yield),
            sd = sd(Yield, na.rm = TRUE),
            n = sum(!is.na(Yield)),
            se = sd / sqrt(n),
            .groups = "drop"
  )


ggplot() +
  geom_jitter(
    data = yield,
    aes(x = Treatment, y = Yield, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = yield_summ,
    aes(x = Treatment, y = mean_yield, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = yield_summ,
    aes(x = Treatment, y = mean_yield, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  
#Global emmean layer
  
geom_line(
  data = yield_emm,
  aes(x = Treatment, y = response, group = 1),
  inherit.aes = FALSE,
  linewidth = 1.6,
  color = "black"
) +
  
  geom_errorbar(
    data = yield_emm,
    aes(x = Treatment,
        ymin = response-SE,
        ymax = response+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = yield_emm,
    aes(x = Treatment, y = response),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Yield (kg/ha)",
  ) +
  theme_minimal(base_size = 14)
```



## Without field 5

```{r Yield dataframe no-5, include=FALSE}
yield_no5 <- yield %>% 
  filter(Field != "5")

model.yield.no5 <- glmmTMB(Yield ~ Treatment + (1|Field), data = yield_no5, family = Gamma (link = "log"))
```

**Summary:**

```{r Yield summary no-5, echo = FALSE}
summary(model.yield.no5)

r2(model.yield.no5)

Anova(model.yield.no5)
```

**Model diagnostics:**

```{r Yield model diagnostics no-5, echo=FALSE}
dharma <- simulateResiduals(model.yield.no5, plot = T)
```

**Estimated marginal means plots:**

```{r Yield emmeans no-5, echo = FALSE, message = FALSE}
emmeans(model.yield.no5, pairwise ~ Treatment, type = "response")

yield.no5_emm <- as.data.frame(emmeans(model.yield.no5, pairwise ~Treatment, type = "response")$emmeans)
ggplot(yield.no5_emm, aes(x = Treatment, y = response))+
         geom_point(size = 2)+
         geom_errorbar(aes(ymin = response-SE, ymax = response+SE, width = 0.2))+
  labs(title = "Estimated Marginal Means of Yield by Treatment",
       x = "Treatment",
       y = "Yield (kg/ha)") +
  theme_minimal()
```

**Figure:**

```{r Figure yield no-5, echo = FALSE}
yield.no5_summ <- yield_no5 %>% 
  group_by(Field, Treatment) %>% 
  summarise(mean_yield = mean(Yield),
            sd = sd(Yield, na.rm = TRUE),
            n = sum(!is.na(Yield)),
            se = sd / sqrt(n),
            .groups = "drop"
  )


ggplot() +
  geom_jitter(
    data = yield_no5,
    aes(x = Treatment, y = Yield, color = as.factor(Field)),
    width = 0.07, height = 0,
    alpha = 0.30, size = 2.5
  ) +
  geom_line(
    data = yield.no5_summ,
    aes(x = Treatment, y = mean_yield, group = as.factor(Field), color = as.factor(Field)),
    linewidth = 0.9, alpha = 0.5
  ) +
  geom_point(
    data = yield.no5_summ,
    aes(x = Treatment, y = mean_yield, color = as.factor(Field)),
    size = 4,
    alpha = 0.5
  ) +
  
  
#Global emmean layer
  
  geom_line(
    data = yield.no5_emm,
    aes(x = Treatment, y = response, group = 1),
    inherit.aes = FALSE,
    linewidth = 1.6,
    color = "black"
  ) +
  
  geom_errorbar(
    data = yield.no5_emm,
    aes(x = Treatment,
        ymin = response-SE,
        ymax = response+SE,
        group = 1),
    inherit.aes = FALSE,
    width = 0.08,
    linewidth = 1.1,
    color = "black"
  ) +
  
  geom_point(
    data = yield.no5_emm,
    aes(x = Treatment, y = response),
    inherit.aes = FALSE,
    size = 4.8,
    shape = 21,
    fill = "white",
    color = "black",
    stroke = 1.2
  ) +
  scale_x_discrete(labels = c("BE" = "No", "FO" = "Yes")) +
  scale_color_manual(values = pal_cb, name = "Field") +
  labs(
    x = "Birds",
    y = "Yield (kg/ha)",
  ) +
  theme_minimal(base_size = 14)
```
